A Switching Planner for Combined Task and Observation Planning

### 2011-02-05 subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
Few small changes

We describe our approach that switches between sequential and
contingent sessions. As a continual planning approach, it proceeds by
interleaving planning and execution in a deterministic-action POMDP
described in DTPDDL. During a sequential session, planning is
performed by a ``classical'' system,\footnote{That is, a planner
designed to solve fully observable deterministic tasks.}  and
execution proceeds according to the {\em trace} computed by that
system. Taking the form of a classical plan, the trace specifies a
sequence of POMDP actions that achieve the agent's objectives in a
deterministic approximation, i.e., {\em determinisation}, of the
problem at hand. More precisely, the trace is an interleaved sequence
of POMDP actions and {\em assumptive} actions. The latter correspond
to assumptions the planner makes about the truth value of propositions
-- e.g. that a box of cornflakes is located in the kitchen at the
third plan step. They are called {\em applicability} assumptions if
the trace includes an action $\action$ that is not applicable with
probability $1$ at the belief-state
\bstate\ that the system is projected to be in when \action\ is
scheduled for execution, i.e., $\exists\state\in\states\;
\bstate(\state) > 0$ and $\poss(\action)\not\subseteq\state$. By
scheduling
\action,  the serial planner makes an assumption about the
observability of the precondition $\poss(\action)$.

Our approach always begins with a sequential session. Non-assumptive
actions from the trace are executed in sequence until the
applicability of the next scheduled action is uncertain. We denote
that action by \switchAction.  A contingent session then begins that
tailors sensory processing by the agent to determine whether the
assumptions made in the trace hold. 
%%
We add {\em disconfirm} and {\em confirm} actions to the model, so
the session is encouraged to judge the assumptions made in the
trace. On execution of one of those actions, control is returned to
the sequential session that continues at the current%%{\em underlying}
belief-state.
%%
%% For each assumption we add a {\em
%% disconfirm} action to the POMDP whose execution is rewarding
%% (resp. costly) if the assumption is false (resp. true). We also add
%% one {\em confirm} action that is rewarding (resp. costly) to execute
%% if $\poss(\switchAction)$ is true (resp. false). If execution of the
%% contingent plan applies a disconfirm action, then a new sequential
%% session begins at the {\em underlying} belief-state. If the confirm
%% action is executed, the sequential session is resumed, and
%% \switchAction\ is executed. Otherwise, control rests in the continual
%% session.
%%
Because contingent planning is only practical in relatively small
POMDPs, contingent sessions plan in an abstract decision process
determined by the current trace. This abstraction is constructed by
first omitting all propositions that do not feature in the trace, and
by then iteratively refining the model while the result is of a
practicable size.

Finally, whether proceeding in a sequential or contingent session, our
continual planner maintains a factored representation of successive
belief-states by performing belief revision according to
Eq~\ref{eq:revision}. Moreover, our internal representation of the
underlying belief-state corresponds more-or-less directly to an
$(\pp{:init})$ declaration. Our approach uses that distribution: (1)
as the source of candidate determinisations for sequential planning,
(2) to determine when to switch, and (3) as a mechanism to guide
construction of an abstract process for contingent sessions.

%% PLEASE STICK A PARAGRAPH HERE, (SENSIBLY) REQUESTED BY MORITZ, ABOUT
%% HOW WE DO BELIEF REVISION ACROSS SWITCHING.

%% \Omit{
%%   Therefore, although our current implementation does not
%%   explicitly support it, our approach generalises to POMDPs with
%%   stochastic action models.
%% }

>>>>

We now describe our {\em switching} planning system that operates
according to the continual planning paradigm, given descriptions of
the current problem and domain in DTPDDL. Recapping, the system {\em
switches}, in the sense that planning proceeds in interleaved
sessions, in which the base planner is either {\em sequential}/linear
or {\em decision-theoretic}.
%%
During a sequential session, a rewarding {\em trace} of a possible
execution is computed using a cost-optimising satisficing planner
which trades action costs, goal rewards, and determinacy.
%%
Formatted as a linear plan, the trace specifies a sequence of actions
that achieves some reachable goals following a deterministic
approximation of the problem at hand.
%%
Structurally, it is a sequence of elements that are either: (i) ground
actions from the DTPDDL description of the world, or (ii) atomic {\em
assumptions}, modelled as deterministic actions, made about the truth
value of facts that can only be determined at runtime (e.g., that a
box of cornflakes is located on the corner bench in the kitchen).
%%
The system always begins in a sequential session, so that plan
execution proceeds by applying DTPDDL actions from the trace in
sequence until the applicability of the next scheduled action is too
uncertain according to a threshold parameter~(here, 95\%). In that
eventuality a DT session commences, which tailors sensory processing
to determine whether the assumptions made in the trace hold, or which
otherwise acts to achieve the goals.

Because online DT planning in large problems is impractical (we seek
response times in seconds), DT sessions plan in an abstract process
determined by the current trace and underlying belief-state. This
abstraction is constructed by first excluding all propositions (and
related actions) that are not true of any state in the trace, then
adding them back, using as a heuristic the entropy of the trace
assumptions conditional on a candidate proposition. Propositions are
added, one at a time, until the number of states in the initial
belief-state reaches a given threshold.\footnote{Our experiments are
performed using a number of different threshold parameters.}  To the
resulting abstract model we also add {\em disconfirm} and {\em
confirm} actions that the contingent session can schedule in order to
judge an atomic assumption in the trace. In the abstract model these
actions yield a relatively small reward if the corresponding judgement
is true (or small penalty otherwise). If a judgement action is
scheduled for execution the contingent session is terminated, and
control is returned to a sequential session.

Finally, whether proceeding in a sequential or contingent session, our
continual planner maintains a factored representation of successive
belief-states by performing belief revision. Our internal
representation of the underlying belief-state corresponds more-or-less
directly to an $(\pp{:init})$ declaration. That belief-state
distribution is used: (1) as the source of candidate determinisations
for sequential planning, (2) in determining when to switch to a DT
session, and (3) as a mechanism to guide construction of an abstract
process for contingent sessions.
